A sample workflow

### 2025-04-23 analyses/african_mammals_workflow.qmd
slowly continuing workflow example, but LPI is a mess

@@

@@

Now we check how well our model predicts our data ...

>>>>

lpi_subset <- lpi_subset[, c("ID", "Genus", "Species", "Units", paste0("X", 1950:2020))]
lpi_subset$genus_species <- paste(lpi_subset$Genus, lpi_subset$Species)

lpi_subset_rsh <- 
  reshape(lpi_subset, direction = "long", idvar = "ID",
          varying =  paste0("X", 1950:2020), v.names = c("value"), timevar = "year", times = 1950:2020)
lpi_subset_rsh <- lpi_subset_rsh[lpi_subset_rsh$value != "NULL",]

@@

```{r}

lpi_subset <- lpi_subset[, c("ID", "Genus", "Species", "Units", paste0("X", 1950:2020))]
lpi_subset$genus_species <- paste(lpi_subset$Genus, lpi_subset$Species)

lpi_subset_rsh <- 
  reshape(lpi_subset, direction = "long", idvar = "ID",
          varying =  paste0("X", 1950:2020), v.names = c("value"), timevar = "year", times = 1950:2020)
lpi_subset_rsh <- lpi_subset_rsh[lpi_subset_rsh$value != "NULL",]
lpi_subset_rsh$value <- as.numeric(lpi_subset_rsh$value)

units_to_keep <- c("Estimate - entire population", "Estimate - entire population (August)",
                   "Individual", "individuals", "Individuals", "Number of adults", "number of individuals",
                   "Number of individuals", "Number of Individuals", "Number of rhinos", "numbers of indviduals",
                   "population estimate", "Population estimate", "Population estimate (individuals)", 
                   "Population estimates", "Total population size")

lpi_subset_rsh <- lpi_subset_rsh[lpi_subset_rsh$Units %in% units_to_keep,]

ggplot(data = lpi_subset_rsh) +
  geom_line(aes(x = year, y = value, 
                group = paste0(genus_species, ID),
                color = genus_species)) +
  theme_bw()

```{r}

lpi_subset_rsh$year <- lpi_subset_rsh$year-1980
nestlm <- lm(value ~ year * genus_species/ID, data = lpi_subset_rsh)
coefs <- summary(nestlm)$coefficients

species <- unique(lpi_subset_rsh$genus_species)[order(unique(lpi_subset_rsh$genus_species))]

species_intercept <- c(coefs["(Intercept)", "Estimate"], sapply(species[2:length(species)], function(sp){coefs["(Intercept)", "Estimate"] + coefs[paste0("genus_species", sp), "Estimate"]}))
species_intercept_sd <- c(coefs["(Intercept)", "Std. Error"], sapply(species[2:length(species)], function(sp){coefs[paste0("genus_species", sp), "Std. Error"]}))

intercepts <- data.frame(est=species_intercept, est.sd=species_intercept_sd, species)

species_slope <- c(coefs["year", "Estimate"], sapply(species[2:length(species)], function(sp){coefs["year", "Estimate"] + coefs[paste0("year:genus_species", sp), "Estimate"]}))
species_slope_sd <- c(coefs["year", "Std. Error"], sapply(species[2:length(species)], function(sp){coefs[paste0("year:genus_species", sp), "Std. Error"]}))

slopes <- data.frame(est=species_slope, est.sd=species_slope_sd, species)

@@

```{r}
#| fig-height: 10
#| fig-width: 12

retrodata <- data.frame(obs = lpi_subset_rsh$value, pred = predict(nestlm),
                        species = lpi_subset_rsh$genus_species, id = lpi_subset_rsh$ID)

ggplot(data = retrodata) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey") +
  facet_wrap(~ species, scales = "free") +
  geom_point(aes(x = obs, y = pred, group = id, color = as.character(id)),
             size = 0.5) + 
  theme_bw() + theme(legend.position = 'none')
